---
title: "Class Activity 1"
Author: Timothy Lee
output: html_document
---


**Load the libraries tidyr and dplyr**

```{r echo = FALSE, message = FALSE}
library(tidyr)
library(dplyr)
```

**Create a data frame from the swirl-data.csv file called DF1**

```{r echo = FALSE}
DF1 = read.csv("C:/Users/Timothy/Google Drive/TC Stuff/Analytics/HUDK 4050 & 4051 - Learning Analytics/projects/class-activity-1/swirl-data.csv", header = TRUE)
```

**The variables are:**

* **`course_name` - the name of the R course the student attempted**
* **`lesson_name` - the lesson name**
* **`question_number` - the question number attempted correct - whether the question was answered correctly**
* **`attempt` - how many times the student attempted the question**
* **`skipped` - whether the student skipped the question**
* **`datetime` - the date and time the student attempted the question**
* **`hash` - anonymyzed student ID**

**Create a new data frame that only includes the variables `hash`, `lesson_name` and `attempt` called DF2**

```{r}
DF2 = DF1 %>%  select("hash", "lesson_name", "attempt")
#DF2 = data.frame('Hash' = DF1$hash, 'LessonName' = DF1$lesson_name, 'Attempt' = DF1$attempt)
```

**Use the `group_by` function to create a data frame that sums all the attempts for each hash by each lesson_name called DF3**

Is this correct?
```{r}
DF3 = DF2 %>% filter(!is.na(attempt)) %>% group_by(hash, lesson_name) %>% summarise(sumAttempts = sum(attempt))
```

**On a scrap piece of paper draw what you think DF3 would look like if all the lesson names were column names**

Something like:

| Hash          | Basic Building Blocks     | Logic |
| ------------- |:-------------------------:| -----:|
| 2864          | 2                         | 45    |
| 4807          | 3                         | 100   |
| 2864          | 45                        | 251   |


Convert DF3 to this format.

```{r}
DF3spread = DF3 %>% spread(key = lesson_name, value = sumAttempts)
```

**Create a new data frame from DF1 called DF4 that only includes the variables `hash`, `lesson_name` and `correct`**

```{r}
DF4 = DF1 %>% select("hash", "lesson_name", "correct")
```

**Convert the correct variable so that `TRUE` is coded as the number 1 and `FALSE` is coded as 0**
```{r}
codedCorrect = ifelse(DF4$correct == TRUE, yes = 1, no = 0)
#No need for additional ifelse for NA, since ifelse documentation says "Missing values in test give missing values in the result."
#I checked and this appears to work.
DF4 = cbind(DF4, codedCorrect)
```

**Create a new data frame called DF5 that provides a mean score for each student on each course**

```{r}
DF5 = DF4 %>% filter(!is.na(codedCorrect)) %>% group_by(hash, lesson_name) %>% summarise(meanScore = mean(codedCorrect))
```

**Extra credit Convert the datetime variable into month-day-year format and create a new data frame (DF6) that shows the average correct for each day**

```{r}
#Select only the relevant variables
DF6 = DF1 %>% select("correct", "datetime")
#Convert score to numbers - 1 for correct, 0 for incorrect
DF6$correct = ifelse(DF6$correct == TRUE, yes = 1, no = 0)
#Change datetime to proper datetime variables
DF6$datetime = as.POSIXct(DF6$datetime, origin = "1970-01-01")
#Omit time so group_by groups by day and not second
DF6$datetime = lubridate::round_date(DF6$datetime, unit = 'day')

DF6 = DF6 %>% filter(!is.na(correct)) %>%
  group_by(datetime) %>%
  summarise(averageCorrect = mean(correct))
DF6

```

